---
title: "Machine Learning Project"
author: "Jared Cassibba"
date: "January 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(caret)
library(ggplot2)
library(randomForest)
```

#read in training and testing data
```{r}
train <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!",""))
test <- read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!",""))

names(train)
str(train)
summary(train)
```
#this is the outcome we want to predict
```{r}
summary(train$classe)
```
#we want to predict the 'classe' variable using any other variable to predict with
```{r}
inTrain <- createDataPartition(y=train$classe, p=0.6, list=FALSE)
myTrain <- train[inTrain, ]
myTest <- train[-inTrain, ]
dim(myTrain)

dim(myTest)
```
#some exploratory plots
```{r}
featurePlot(x=train[, 150:159], y = train$classe, plot = 'pairs')
```
#first we will remove variables with mostly NAs (use threshold of >75%)
```{r}
mytrain_SUB <- myTrain
for (i in 1:length(myTrain)) {
  if (sum(is.na(myTrain[ , i])) / nrow(myTrain) >= .75) {
    for (j in 1:length(mytrain_SUB)) {
      if (length(grep(names(myTrain[i]), names(mytrain_SUB)[j]))==1) {
        mytrain_SUB <- mytrain_SUB[ , -j]
      }
    }
  }
}

dim(mytrain_SUB)
names(mytrain_SUB)
```
#remove columns that are obviously not predictors
```{r}
mytrain_SUB2 <- mytrain_SUB[,8:length(mytrain_SUB)]
```
#remove variables with near zero variance
```{r}
NZV <- nearZeroVar(mytrain_SUB2, saveMetrics = TRUE)

keep <- names(mytrain_SUB2)
```

#Random Forest Model
#fit model- RANDOM FOREST
```{r}
set.seed(3)
modFit <- randomForest(classe~., data = mytrain_SUB2)
print(modFit)
```
#cross validation on my testing data
#out of sample error
```{r}
predict1 <- predict(modFit, myTest, type = "class")
confusionMatrix(myTest$classe, predict1)
```
#in sample error
```{r}
predict_train <- predict(modFit, myTrain, type = "class")
confusionMatrix(myTrain$classe, predict_train)

predict_FINAL <- predict(modFit, test, type = "class")
print(predict_FINAL)
```